---
title: "Time_Series_Project"
author: "Yash Bajaj"
date: "4/19/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(astsa)
library(dplyr)
library(faraway)
library(lmtest)
library(MASS)
library(forecast)
library(car)
library(fastDummies)
library(Metrics)
library(leaps)
```


```{r}
df <- read.csv('C:/Users/hp/Documents/R/STAT 429/Project/Behavior of the urban traffic of the city of Sao Paulo in Brazil/Behavior of the urban traffic of the city of Sao Paulo in Brazil.csv',sep=';')
df
```

```{r}
df$Slowness.in.traffic....<-as.numeric(gsub(",", ".", gsub("\\.", "", df$Slowness.in.traffic....)))
```

```{r}
df
```

```{r}
ts.plot(diff(df$Slowness.in.traffic....))
```



Monday - Friday
```{r}
a1 <- length(df$Slowness.in.traffic....)
b1 <- c()
weekday <- c('Monday','Tuesday','Wednesday','Thursday','Friday')
for(i in 0:(a1/27 -1)){
      b1[(27*i+1):(27*i+28)] <- weekday[((i%%length(weekday))+1)]
  }
df$weekday <- b1[1:length(df$Slowness.in.traffic....)]
```

```{r}
df
```

```{r}
data <- df
data$weekday <- as.factor(data$weekday)
data$Hour..Coded. <- as.factor(data$Hour..Coded.)
data$Time <- 1:135
```


```{r}
data
```

```{r}
summary(lm(Slowness.in.traffic....~Time,data=data))
```

```{r}
plot(data$Time,data$Slowness.in.traffic....)
```

```{r}
linear <- lm(Slowness.in.traffic....~.,data=data)
summary(linear)
```
```{r}
linear1 <- lm(Slowness.in.traffic....~.-Time,data=data)

x = model.matrix(linear1)[,-1] 
x = x - matrix(apply(x,2, mean), dim(x)[1],dim(x)[2], byrow=TRUE)
x = x / matrix(apply(x, 2, sd), dim(x)[1],dim(x)[2], byrow=TRUE)
#Extracting the eigen-values:
e = eigen(t(x) %*% x) 
sqrt(max(e$val)/min(e$val))
```

```{r}
vif(linear1)
```

#Leverage points
```{r}
n=dim(data)[1]; # sample size
p=19; # predictors plus intercept

# Compute Leverages
lev=influence(linear1)$hat

# Determine which exceed the 2p/n threshold
newlev = lev[lev>2*p/n]
length(newlev)
```

```{r}
length(newlev)/n
```

```{r}
# Prepare a half-normal plot 
halfnorm(lev, ylab="Leverages")
```

#Checking presence of outliers-
```{r}
jackknife=rstudent(linear1)
n = dim(data)[1]
p = 19
x=qt(.05/(2*n), n-p-1) #Significance level adjusted with Bonferroni's Correction
x
```
```{r}
sort(abs(jackknife),decreasing =T)[1:10]
```
#Checking Highly influential points -
```{r}
cook = cooks.distance(linear1)
# Extract max Cook's Distance
max(cook,na.rm=TRUE)
```

```{r}
halfnorm(cook, 6, labs=as.character(1:length(cook)), ylab="Cook's distances")
```

Since cook's distance is less than 1, we don't remove any data points.

#Checking for normality assumptions -
```{r}
plot(linear1,which=1)
```

Since there is no trend in our residual vs fitted values plot, we see
that normality assumptions hold true.

#Breush Pagan Test-
```{r}
bptest(linear1)
```
Since p-value of the test is greater than our significance level of 0.05 , we fail to reject H0 and conclude that the error variance is constant.

#Checking the normality of errors assumption-
Let’s check if the errors in our model are normally distributed around a mean of zero. For that, we need to look at the histogram of the residuals as well as the Q-Q plot .
```{r}
hist(linear1$residuals, breaks = 10)
```

```{r}
plot(linear1,which = 2)
```

#Shapiro-Wilks test-
```{r}
shapiro.test(linear1$residuals)
```
#Box-Cox -
```{r}
box <- boxcox(Slowness.in.traffic....~.,data=data)
(lambda <- bc$x[which.max(bc$y)])
```

```{r}
tran_model <- lm(((Slowness.in.traffic....^lambda-1)/lambda) ~ .,data=data)
```

```{r}
shapiro.test(tran_model$residuals)
```

#Checking the linearity assumption-
```{r}
avPlots(linear)
```

Removing time as it has perfect multicollinearity with weekend.
```{r}
drop <- c('Time')
data.new = data[,!(names(data) %in% drop)]
names(data.new)
```


```{r}
full.model <- lm(Slowness.in.traffic....~.,data=data.new)
summary(full.model)
```
```{r}
checkresiduals(full.model)
```


```{r}
AIC(full.model)
BIC(full.model)
```

#RMSE-

## RMSE Func

```{r}
full = predict(full.model, newdata=data.new[,-18])
calc_rmse<-function(x,y){
  
  diff=x-y
  
  sqrt(mean(diff^2))
}

calc_rmse(full,data.new$Slowness.in.traffic....)
```

#Step function using direction ='both'
```{r}
step_both_linear <- step(full.model,direction='both')
```

#Fitting model using step=both -
```{r}
step_both_linear_fit <- lm(Slowness.in.traffic.... ~ Hour..Coded. + Accident.victim + Occurrence.involving.freight + Lack.of.electricity + Point.of.flooding + Defect.in.the.network.of.trolleybuses + 
Tree.on.the.road + Intermittent.Semaphore + weekday,data = data.new)
summary(step_both_linear_fit)
```

```{r}
checkresiduals(step_both_linear_fit)
```

```{r}
AIC(step_both_linear_fit)
BIC(step_both_linear_fit)
```

#RMSE-
```{r}
step_both = predict(step_both_linear_fit, newdata=data.new[,-18])
calc_rmse<-function(x,y){
  
  diff=x-y
  
  sqrt(mean(diff^2))
}

calc_rmse(step_both,data.new$Slowness.in.traffic....)
```
#Step function direction=forward -
```{r}
step_forward_linear <- step(full.model,direction='forward')
```

#Fitting model using step=forward -
```{r}
step_forward_linear_fit <- lm(Slowness.in.traffic.... ~ Hour..Coded. +              Immobilized.bus + Broken.Truck + Vehicle.excess + Accident.victim +           Running.over + Fire.vehicles + 
    Occurrence.involving.freight + Incident.involving.dangerous.freight + 
    Lack.of.electricity + Fire + Point.of.flooding + Manifestations + 
    Defect.in.the.network.of.trolleybuses + Tree.on.the.road + 
    Semaphore.off + Intermittent.Semaphore + weekday,data = data.new)
summary(step_forward_linear_fit)
```

```{r}
checkresiduals(step_forward_linear_fit)
```

```{r}
AIC(step_forward_linear_fit)
BIC(step_forward_linear_fit)
```

#RMSE-
```{r}
step_forward = predict(step_forward_linear_fit, newdata=data.new[,-18])
calc_rmse<-function(x,y){
  
  diff=x-y
  
  sqrt(mean(diff^2))
}

calc_rmse(step_forward,data.new$Slowness.in.traffic....)
```

#Step function direction=backward -
```{r}
step_backward_linear <- step(full.model,direction='backward')
```

#Fitting model using step=backward -
```{r}
step_backward_linear_fit <- lm(Slowness.in.traffic.... ~Hour..Coded. + Accident.victim + Occurrence.involving.freight + 
    Lack.of.electricity + Point.of.flooding + Defect.in.the.network.of.trolleybuses + 
    Tree.on.the.road + Intermittent.Semaphore + weekday,data = data.new)
summary(step_backward_linear_fit)
```

```{r}
checkresiduals(step_backward_linear_fit)
```

```{r}
AIC(step_backward_linear_fit)
BIC(step_backward_linear_fit)
```

#RMSE-
```{r}
step_backward = predict(step_backward_linear_fit, newdata=data.new[,-18])
calc_rmse<-function(x,y){
  
  diff=x-y
  
  sqrt(mean(diff^2))
}

calc_rmse(step_backward,data.new$Slowness.in.traffic....)
```

#Leaps
```{r}
leap = regsubsets(Slowness.in.traffic.... ~ ., data=data.new, nvmax = ncol(data.new)-1)
leap.summary = summary(leap)

n = dim(data.new)[1]
msize = 2:ncol(data.new)

par(mfrow=c(2,2))
plot(msize, leap.summary$adjr2, xlab="No. of Parameters", ylab = "Adjusted Rsquare") 
Aic = n*log(leap.summary$rss/n) + 2*msize; 
plot(msize, Aic, xlab="No. of Parameters", ylab = "AIC"); 
Bic = n*log(leap.summary$rss/n) + msize*log(n); 
plot(msize, Bic, xlab="No. of Parameters", ylab = "BIC")
plot(msize, leap.summary$cp, xlab="No. of Parameters", ylab = "Mallow's Cp")
```

```{r}
rs<-leap.summary
colnames(rs$which)[rs$which[which.max(rs$adjr2),]][-1]
```
```{r}
adjr2.model <- lm(Slowness.in.traffic....~  Hour..Coded.        
+ Lack.of.electricity                  
+ Point.of.flooding                 
+ Defect.in.the.network.of.trolleybuses
+ Tree.on.the.road                   
+ weekday,data = data.new)
summary(adjr2.model)
```

```{r}
checkresiduals(adjr2.model)
```

```{r}
AIC(adjr2.model)
BIC(adjr2.model)
```

#RMSE-
```{r}
leap_adjr2 = predict(adjr2.model, newdata=data.new[,-18])
calc_rmse<-function(x,y){
  
  diff=x-y
  
  sqrt(mean(diff^2))
}

calc_rmse(leap_adjr2,data.new$Slowness.in.traffic....)
```


```{r}
col = colnames(rs$which)[rs$which[which.min(Bic),]][-1]
col
```


```{r}
bic.model <- lm(Slowness.in.traffic....~  Hour..Coded.        
+ Lack.of.electricity                  
+ Point.of.flooding                 
+ Defect.in.the.network.of.trolleybuses
+ Intermittent.Semaphore                   
+ weekday,data = data.new)
summary(bic.model)
```

```{r}
checkresiduals(bic.model)
```

```{r}
AIC(bic.model)
BIC(bic.model)
```

#RMSE-
```{r}
leap_bic = predict(bic.model, newdata=data.new[,-18])
calc_rmse<-function(x,y){
  
  diff=x-y
  
  sqrt(mean(diff^2))
}

calc_rmse(leap_bic,data.new$Slowness.in.traffic....)
```


(Intercept)                           6.89e-08 ***
Hour..Coded.2                         0.977757    
Hour..Coded.3                         0.252036    
Hour..Coded.4                         0.089922 .  
Hour..Coded.5                         0.001013 ** 
Hour..Coded.6                         0.009653 ** 
Hour..Coded.7                         0.072726 .  
Hour..Coded.8                         0.058399 .  
Hour..Coded.9                         0.028195 *  
Hour..Coded.10                        0.025387 *  
Hour..Coded.11                        0.034785 *  
Hour..Coded.12                        0.103672    
Hour..Coded.13                        0.015226 *  
Hour..Coded.14                        0.037433 *  
Hour..Coded.15                        0.130840    
Hour..Coded.16                        0.017217 *  
Hour..Coded.17                        0.009853 ** 
Hour..Coded.18                        0.007725 ** 
Hour..Coded.19                        0.000458 ***
Hour..Coded.20                        6.49e-05 ***
Hour..Coded.21                        3.50e-06 ***
Hour..Coded.22                        1.02e-08 ***
Hour..Coded.23                        3.49e-10 ***
Hour..Coded.24                        3.28e-10 ***
Hour..Coded.25                        2.80e-12 ***
Hour..Coded.26                        1.63e-11 ***
Hour..Coded.27                        2.88e-06 ***
Accident.victim                       0.178102    
Occurrence.involving.freight          0.155181    
Lack.of.electricity                   0.001445 ** 
Point.of.flooding                     3.48e-05 ***
Defect.in.the.network.of.trolleybuses 0.117714    
Tree.on.the.road                      0.035118 *  
Intermittent.Semaphore                0.194071    
weekdayMonday                         6.16e-10 ***
weekdayThursday                       0.051972 .  
weekdayTuesday                        0.002761 ** 
weekdayWednesday                      0.010934 *

```{r}
final_data <- subset(data.new,select = c('Hour..Coded.','Lack.of.electricity','Point.of.flooding','Tree.on.the.road','weekday','Slowness.in.traffic....'))
final_data
```

```{r}
final_data_fac <- dummy_cols(final_data, select_columns = c('Hour..Coded.','weekday'),remove_first_dummy = TRUE,remove_selected_columns = TRUE)
final_data_fac = subset(final_data_fac,select=c('Hour..Coded._5','Hour..Coded._19','Hour..Coded._20','Hour..Coded._21','Hour..Coded._22','Hour..Coded._23','Hour..Coded._24','Hour..Coded._25','Hour..Coded._26','Hour..Coded._27','Lack.of.electricity','Point.of.flooding','Tree.on.the.road','weekday_Monday','Slowness.in.traffic....'))
  final_data_fac
```

```{r}
final_model <- lm(Slowness.in.traffic....~.,data=final_data_fac)
summary(final_model)
```
'Hour..Coded._9','Hour..Coded._10','Hour..Coded._11','Hour..Coded._13','Hour..Coded._14','Hour..Coded._16','Hour..Coded._17','Hour..Coded._18',,'weekday_Wednesday',,'weekday_Tuesday',


```{r}
checkresiduals(final_model)
```

```{r}
AIC(final_model)
BIC(final_model)
```

#RMSE-
```{r}
final_rmse = predict(final_model, newdata=final_data_fac[,-15])
calc_rmse<-function(x,y){
  
  diff=x-y
  
  sqrt(mean(diff^2))
}

calc_rmse(final_rmse,final_data_fac$Slowness.in.traffic....)
```
```{r}
acf2(final_model$residuals)
```
```{r}
auto.arima(final_model$residuals)
library(tseries)
adf.test(final_model$residuals)
tsplot(final_model$residuals)

box.cox(final_model$residuals)

res_ser<-final_model$residuals
tsplot(log(final_model$residuals))

lambda<-BoxCox.lambda(res_ser)

plot.ts(BoxCox(res_ser, lambda = lambda))

res_ser.tr<-BoxCox(res_ser, lambda = lambda)

acf2(res_ser.tr)

auto.arima(res_ser.tr)
```
```{r}

design<-model.matrix(final_model)[,-1]
sarima(res_ser.tr,1,0,0,xreg = design)
```



```{r}
fit <- nnetar(data.new$Slowness.in.traffic...., lambda=0)
autoplot(forecast(fit,h=135))
```



# ``````{r}
# trans_df<-function(df1){
# df1$date <- as.Date(df1$date)
# df1$date<-as.POSIXct(df1$date)
# 
# df1$month<-factor(format(df1$date,"%m"))
# df1$year<-as.numeric(format(df1$date,"%Y"))
# df1$day<-factor(format(df1$date,"%d"))
# df1$wday <- factor(weekdays(as.Date(df1$date)))
# df1$store<-factor(df1$store)
# 
# df1$item<-factor(df1$item)
# df1<-df1[,!colnames(df1) %in% c('date','id')]
# }
# 
# ```
# ```{r}
# df1<-trans_df(df1)

```